
library(data.table)
library(ggplot2)
library(tidyr)

data <- fread("QVI_data_task2.csv")
data <- fread("QVI_transaction.csv")
view(data)

## Select control stores
#create the metrics of interest and filter to stores that are present throughout the pre-trial period.
# Add a new month ID column in the data with the format yyyymmdata.[, YEARMONTH := ]
```{r}
#Create a month ID 
data[, YEARMONTH := year(DATE)*100 + month(DATE)]
```

#### Next, we define the measure calculations to use during the analysis.
```{r}
#### Define the measure calculations 
measureOverTime <- data[, .(totSales = sum(TOT_SALES),
                            nCustomers = uniqueN(LYLTY_CARD_NBR), 
                            nTxnPerCust = uniqueN(TXN_ID)/uniqueN(LYLTY_CARD_NBR), 
                            nChipsPerTxn = sum(PROD_QTY)/uniqueN(TXN_ID),  
                            avgPricePerUnit = sum(TOT_SALES)/sum(PROD_QTY)),
                        by=c("STORE_NBR","YEARMONTH")][order(STORE_NBR, YEARMONTH)]
```

```{r}
####Filter to the pre‐trial period and stores with full observation periods 
storesWithFullObs <- unique(measureOverTime[,.N,STORE_NBR][N== 12, STORE_NBR]) 
preTrialMeasures <- measureOverTime[YEARMONTH< 201902 & STORE_NBR %in% storesWithFullObs,]
```


### We can calculate how correlated the performance of each store is to the trial store
```{r}
####Create a function to calculate correlation for a measure
####looping through each control store
calculateCorrelation <- function(inputTable,metricCol,storeComparison){ 
calcCorrTable= data.table(Store1= numeric(),Store2= numeric(),corr_measure = numeric()) 

    storeNumbers <- unique(inputTable[,STORE_NBR])

    for (i in storeNumbers){ calculatedMeasure = data.table("Store1"=storeComparison,"Store2"=i,     "corr_measure"= cor(inputTable[STORE_NBR== storeComparison,  eval(metricCol)],inputTable[STORE_NBR== i, eval(metricCol)]))

    calcCorrTable <- rbind(calcCorrTable,calculatedMeasure) 
    }
    return(calcCorrTable) 
}
```


## calculate a standardised  metric based on the absolute difference between the trial store’s performance and each control store’s performance.

```{r}
####Create a function to calculate a standardised magnitude distance for a measure,
####looping through each control store
calculateMagnitudeDistance <- function(inputTable,metricCol,storeComparison)
  { calcDistTable= data.table(Store1= numeric(),Store2= numeric(),YEARMONTH= numeric(),measure= numeric()) 
  storeNumbers <- unique(inputTable[,STORE_NBR])
  for (i in storeNumbers)
  {
  calculatedMeasure= data.table("Store1"=storeComparison ,"Store2"=i,
                                "YEARMONTH"=inputTable[STORE_NBR== storeComparison,YEARMONTH],
                                "measure"= abs(inputTable[STORE_NBR==storeComparison,eval(metricCol)] - inputTable[STORE_NBR== i, eval(metricCol)]) 
  ) 
calcDistTable <- rbind(calcDistTable,calculatedMeasure) 
  }

####Standardise the magnitude distance so that the measure ranges from 0 to 1  
  minMaxDist <- calcDistTable[,.(minDist= min(measure),maxDist= max(measure)),
                            by= c("Store1","YEARMONTH")] 
  distTable <- merge(calcDistTable,minMaxDist,by= c("Store1","YEARMONTH"))
  distTable[,magnitudeMeasure:= 1 - (measure - minDist)/(maxDist - minDist)] 
  finalDistTable <- distTable[,.(mag_measure= mean(magnitudeMeasure)),by= .(Store1,Store2)] 
  return(finalDistTable) 
}
```



### Now let’s use the functions to find the control stores!
#select control stores based on how similar monthly total sales in dollar amount and monthly number of customers are to the trial stores. So we will need to use our functions to get four scores, two for each of total sales and total customers.
```{r}
#### Usethe functions for calculating correlation 
trial_store <- 77 
corr_nSales <- calculateCorrelation(preTrialMeasures, quote(totSales), trial_store) 
corr_nCustomers <- calculateCorrelation(preTrialMeasures,quote(nCustomers), trial_store) 
####Usethe functions for calculating magnitude
magnitude_nSales <- calculateMagnitudeDistance(preTrialMeasures, quote(totSales),trial_store) 
magnitude_nCustomers <- calculateMagnitudeDistance(preTrialMeasures,  quote(nCustomers),trial_store)
```

####Usethe functions for calculating magnitude
```{r}
magnitude_nSales <- calculateMagnitudeDistance(preTrialMeasures, quote(totSales),trial_store) 
magnitude_nCustomers <-calculateMagnitudeDistance(preTrialMeasures, quote(nCustomers),trial_store)

```

####Create a combined score composed of correlation and magnitude 
```{r}
corr_weight <- 0.5
score_nSales <- merge(corr_nSales,magnitude_nSales,by= c("Store1", "Store2"))[,scoreNSales := corr_measure * corr_weight + mag_measure * (1-corr_weight)] 
score_nCustomers <- merge(corr_nCustomers,magnitude_nCustomers,by= c("Store1","Store2"))[,scoreNCust := corr_measure * corr_weight + mag_measure * (1-corr_weight)]
```

####Combine scores across the drivers 
```{r}
score_Control <- merge(score_nSales,score_nCustomers,by= c("Store1", "Store2")) 
score_Control[,finalControlScore:= scoreNSales * 0.5 + scoreNCust * 0.5]
```

####Select control stores based on the highest matching store(closest to 1 but not the store itself, i.e. the second ranked highest store)
####Select control store for trial store 77 

```{r}
control_store <- score_Control[Store1== trial_store, ][order(-finalControlScore)][2,Store2] 
control_store
```


```{r}
####Visual checks on trends based on the drivers 
measureOverTimeSales <- measureOverTime 
pastSales <- measureOverTimeSales[,Store_type:= ifelse(STORE_NBR == trial_store,"Trial", 
                                                       ifelse(STORE_NBR == control_store,
                                                              "Control","Otherstores")) 
                                  ][,totSales:=mean(totSales),by=c("YEARMONTH", "Store_type") 
                                    ][,TransactionMonth:=as.Date(paste(YEARMONTH%/% 100,YEARMONTH %% 100,1,sep="‐"),"%Y‐%m‐%d")        ][YEARMONTH < 201903,] 

ggplot(pastSales, aes(TransactionMonth,totSales,color=Store_type)) + geom_line() + labs(x= "Monthofoperation",y= "Totalsales",title="Totalsalesby month")
```


###Next, number of customers
```{r}
####Visual checks on trends based on the drivers
measureOverTimeCusts <- measureOverTime 
pastCustomers <- measureOverTimeCusts[,Store_type:= ifelse(STORE_NBR == trial_store,"Trial",
                                                    ifelse(STORE_NBR == control_store, "Control","Otherstores")) ][,numberCustomers := mean(nCustomers),by= c("YEARMONTH","Store_type")][,TransactionMonth:= as.Date(paste (YEARMONTH%/% 100,YEARMONTH %% 100,1,sep="-"),"%Y-%m-%d") ][YEARMONTH < 201903,]

ggplot(pastCustomers, aes(TransactionMonth,numberCustomers,color= Store_type)) +  geom_line() + labs(x="Month of operation",y="Total number of customers", title= "Total number of  customers by month")
```

### Assessment of trial

#### Scale pre‐trial contro lsales to match pre‐tria ltrial store sales
# This code is performing statistical scaling or normalization to make the control store comparable to the trial store before analysing the trial period impact.

```{r}
scalingFactorForControlSales <- preTrialMeasures[STORE_NBR== trial_store & YEARMONTH < 201902, sum(totSales)]/preTrialMeasures[STORE_NBR== control_store & YEARMONTH < 201902, sum(totSales)]
#### Apply the scaling factor 
measureOverTimeSales<- measureOverTime 
scaledControlSales <- measureOverTimeSales[STORE_NBR== control_store,][, controlSales := totSales * scalingFactorForControlSales]

```


```{r}
####Calculate the percentage difference between scaled control sales and trial sales
percentageDiff<- merge(scaledControlSales[,c("YEARMONTH","controlSales")],
measureOverTime[STORE_NBR== trial_store, c("totSales", "YEARMONTH")], 
by= "YEARMONTH" )[,percentageDiff := abs(controlSales - totSales)/controlSales]

```

# Let’s see if the difference is significant!
```{r}
####As our null hypothesis is that the trial period is the same as the pre‐trial period,let's take the standard deviation based on the scaled percentage difference in the pre‐trial period
stdDev <- sd(percentageDiff[YEARMONTH< 201902,percentageDiff]) 

####note that there are 8 months in the pre‐trial period
####hence 8‐1 =7 degrees of freedom
degreesOfFreedom<-7  

####We will test with a null hypothesis of there being 0 difference between trial 
####and control stores
percentageDiff[,tValue:= (percentageDiff - 0)/stdDev ][,TransactionMonth :=  as.Date(paste(YEARMONTH %/% 100, YEARMONTH %% 100,1, sep="-"),"%Y-%m-%d") 
                                                       ][YEARMONTH < 201905 & YEARMONTH > 201901, .(TransactionMonth, tValue)]

```

####Find the 95th percentile of the t distribution with the appropriate 
####degrees of freedom to compare against 

```{r}
qt(0.95,df= degreesOfFreedom)
```

```{r}
measureOverTimeSales<-measureOverTime 
####Trial and control store total sales 
pastSales<- measureOverTimeSales[,Store_type:= ifelse(STORE_NBR == trial_store,"Trial", 
                                                      ifelse(STORE_NBR == control_store, "Control","Otherstores"))][,totSales:=mean(totSales),by=c("YEARMONTH", "Store_type") 
                                                                                                                    ][,TransactionMonth:=as.Date(paste(YEARMONTH%/% 100,YEARMONTH %% 100,1,sep="-"),"%Y-%m-%d")                                                                                          ][Store_type %in% c("Trial","Control"),]

####Control store 95th percentile 
pastSales_Controls95 <- pastSales[Store_type== "Control", ][,totSales := totSales * (1 + stdDev * 2) ][,Store_type := "Control 95th% confidence interval"] 

####Control store 5th percentile 
pastSales_Controls5 <- pastSales[Store_type== "Control", ][,totSales := totSales * (1 - stdDev * 2) ][,Store_type := "Control 5th% confidence interval"]

trialAssessment <- rbind(pastSales,pastSales_Controls95,pastSales_Controls5)
####Plotting these in one nice graph

ggplot(trialAssessment,aes(TransactionMonth,totSales,color=Store_type)) + geom_rect(data=trialAssessment[YEARMONTH< 201905 & YEARMONTH > 201901,],                  aes(xmin= min(TransactionMonth),xmax= max(TransactionMonth), ymin=0, ymax=Inf,color=NULL),show.legend=FALSE) + geom_line() + labs(x= "Month of operation", 
y= "Total sales", title="Total sales by month")
```

#Repeat finding the control store and assessing the impact of the trial for each of the other two trial stores.
#Trial store 86
```{r}
measureOverTime <- data[,.(totSales= sum(TOT_SALES), 
                           nCustomers= uniqueN(LYLTY_CARD_NBR), 
                           nTxnPerCust= uniqueN(TXN_ID)/uniqueN(LYLTY_CARD_NBR), 
                           nChipsPerTxn= sum(PROD_QTY)/uniqueN(TXN_ID), 
                           avgPricePerUnit= sum(TOT_SALES)/sum(PROD_QTY) ) ,by=c("STORE_NBR","YEARMONTH")][order(STORE_NBR, YEARMONTH)] 

#### Use the functions for calculating correlation 
trial_store <- 86 
corr_nSales <- calculateCorrelation(preTrialMeasures,quote(totSales), trial_store) 
corr_nCustomers <- calculateCorrelation(preTrialMeasures,quote(nCustomers), trial_store) 

#### Use the functions for calculating magnitude 
magnitude_nSales <- calculateMagnitudeDistance(preTrialMeasures, quote(totSales),trial_store)
magnitude_nCustomers <- calculateMagnitudeDistance(preTrialMeasures, quote(nCustomers),trial_store) 

###Create a combined score composed of correlation and magnitude 
corr_weight <- 0.5 
score_nSales <- merge(corr_nSales,magnitude_nSales,by= c("Store1", "Store2"))[,scoreNSales := corr_measure * corr_weight + mag_measure * (1-corr_weight)]
score_nCustomers <- merge(corr_nCustomers,magnitude_nCustomers,by= c("Store1","Store2"))[,scoreNCust := corr_measure * corr_weight + mag_measure * (1-corr_weight)] 

####Combine scores across the drivers
score_Control <- merge(score_nSales,score_nCustomers,by= c("Store1", "Store2")) 
score_Control[,finalControlScore:= scoreNSales * 0.5 + scoreNCust * 0.5] 
####Select control stores based on the highest matching store 
####(closest to 1 but not the store itself, i.e.the second ranked highest store)
####Select control store for trial store 86 
control_store <- score_Control[Store1== trial_store, ][order(-finalControlScore)][2,Store2] 
control_store

```

#Looks like store 155 will be a control store for trial store86. 
#Again, let’s check visually if the drivers are indeed similar in the period before the trial.
#look at total sales first
```{r}
####Visual checks on trends based on the drivers
measureOverTimeSales<- measureOverTime 
pastSales <- measureOverTimeSales[,Store_type:= ifelse(STORE_NBR == trial_store,"Trial", 
                                                       ifelse(STORE_NBR == control_store,
                                                              "Control","Otherstores")) 
                                  ][,totSales:=mean(totSales),by=c("YEARMONTH", "Store_type") 
                                    ][,TransactionMonth:=as.Date(paste(YEARMONTH%/% 100,YEARMONTH %% 100,1,sep="-"),"%Y-%m-%d")
                                      ][YEARMONTH < 201903,] 

ggplot(pastSales, aes(TransactionMonth,totSales,color=Store_type)) + geom_line(aes(linetype=Store_type)) + labs(x= "Month of operation",                                                                  y= "Total sales",title = "Total sales by month")

```

#Let’s now assess the impact of the trial on sales.
```{r}
#### Scale pre‐trial contro lsales to match pre‐trial trial store sales
scalingFactorForControlSales <- preTrialMeasures[STORE_NBR== trial_store & YEARMONTH < 201902, sum(totSales)]/preTrialMeasures[STORE_NBR== control_store & YEARMONTH < 201902, sum(totSales)]
#### Apply the scaling factor 
measureOverTimeSales <- measureOverTime 
scaledControlSales <- measureOverTimeSales[STORE_NBR== control_store,][, controlSales := totSales * scalingFactorForControlSales]

####Calculate the percentage difference between scaled control sales and trial sales
percentageDiff <- merge(scaledControlSales[,c("YEARMONTH","controlSales")], measureOverTime[STORE_NBR== trial_store, c("totSales", "YEARMONTH")],                         by= "YEARMONTH" )[,percentageDiff := abs(controlSales-totSales)/controlSales]

####As our null hypothesisis that the trial period is the same as the pre‐trial period, let's take the standard deviation based on the scaled percentage difference in the pre‐trial period
stdDev<- sd(percentageDiff[YEARMONTH< 201902,percentageDiff])
degreesOfFreedom<- 7
####Trial and control store total sales 

measureOverTimeSales<- measureOverTime 
pastSales <- measureOverTimeSales[,Store_type:= ifelse(STORE_NBR == trial_store,"Trial", 
                                                       ifelse(STORE_NBR == control_store, "Control","Otherstores")) 
                                  ][,totSales:=mean(totSales),by=c("YEARMONTH", "Store_type") 
                                    ][,TransactionMonth:=as.Date(paste(YEARMONTH%/% 100,YEARMONTH %% 100,1,sep="-"),"%Y-%m-%d") 
                                      ][Store_type %in% c("Trial","Control"),] 
####Control store 95th percentile 
pastSales_Controls95 <- pastSales[Store_type== "Control", ][,totSales := totSales * (1 + stdDev * 2) ][,Store_type := "Control 95th% confidence interval"] 

####Control store 5th percentile 

pastSales_Controls5 <- pastSales[Store_type== "Control", ][,totSales := totSales * (1-stdDev * 2) ][,Store_type := "Control5th%confidence interval"] 

trialAssessment <- rbind(pastSales,pastSales_Controls95,pastSales_Controls5)

####Plotting these in one nice graph

ggplot(trialAssessment,aes(TransactionMonth,totSales,color=Store_type)) + geom_rect(data=trialAssessment[YEARMONTH< 201905 & YEARMONTH > 201901,],
aes(xmin= min(TransactionMonth),xmax= max(TransactionMonth), ymin=0, ymax=Inf,color=NULL),show.legend=FALSE) +
  geom_line(aes(linetype=Store_type)) + labs(x= "Month of operation",y= "Total sales", title="Total sales by month")
```

# assessing this for number of customers as well
```{r}
### Scale pre‐trial control customers to match pre‐trial trial store customers
scalingFactorForControlCust <- preTrialMeasures[STORE_NBR== trial_store & YEARMONTH < 201902, sum(nCustomers)]/preTrialMeasures[STORE_NBR== control_store & YEARMONTH < 201902, sum(nCustomers)]
#### Apply the scaling factor 
measureOverTimeCusts <- measureOverTime 
scaledControlCustomers <- measureOverTimeCusts[STORE_NBR== control_store,  ][,controlCustomers:=nCustomers * scalingFactorForControlCust 
  ][,Store_type:=ifelse(STORE_NBR == trial_store,"Trial",
                        ifelse(STORE_NBR == control_store, "Control","Otherstores")) ]

####Calculate the percentage difference between scaled control sales and trial sales
percentageDiff <- merge(scaledControlCustomers[,c("YEARMONTH", "controlCustomers")],
                        measureOverTime[STORE_NBR== trial_store, c("nCustomers","YEARMONTH")],by= "YEARMONTH" )[ ,percentageDiff :=  abs(controlCustomers-nCustomers)/controlCustomers]

####As our null hypothesis is that the trial period is the same as the pre-trial period, let's take the standard deviation based on the scaled percentage difference in the pre-trial period
stdDev <- sd(percentageDiff[YEARMONTH< 201902,percentageDiff]) 
degreesOfFreedom<- 7

```

```{r}
#### Trial and control store number of customers 
pastCustomers <- measureOverTimeCusts[,nCusts:= mean(nCustomers),by= c("YEARMONTH","Store_type") 
                                      ][Store_type %in% c("Trial","Control"),] 
####Control store 95th percentile 
pastCustomers_Controls95 <- pastCustomers[Store_type== "Control", ][,nCusts := nCusts * (1 + stdDev * 2) ][,Store_type := "Control 95th% confidence interval"] 
####Control store 5th percentile 
pastCustomers_Controls5 <- pastCustomers[Store_type== "Control", ][,nCusts := nCusts * (1 -stdDev * 2) ][,Store_type := "Control 5th% confidence interval"] 

trialAssessment<- rbind(pastCustomers,pastCustomers_Controls95, pastCustomers_Controls5)

####Plotting these in one nice graph
ggplot(trialAssessment,aes(TransactionMonth,nCusts,color=Store_type)) + geom_rect(data=trialAssessment[YEARMONTH< 201905 & YEARMONTH > 201901,], aes(xmin= min(TransactionMonth),xmax= max(TransactionMonth), ymin=0, ymax=Inf,color=NULL),show.legend=FALSE) + geom_line() + labs(x="Month of operation", y="Total number of customers", title="Total numberof customers by month")
```

#looks like the number of customers is significantly higher in all of the three months

# Trial store88
```{r}
measureOverTime <- data[,.(totSales= sum(TOT_SALES), nCustomers= uniqueN(LYLTY_CARD_NBR), nTxnPerCust= uniqueN(TXN_ID)/uniqueN(LYLTY_CARD_NBR), 
                           nChipsPerTxn= sum(PROD_QTY)/uniqueN(TXN_ID), 
                           avgPricePerUnit= sum(TOT_SALES)/sum(PROD_QTY) ) ,by=c("STORE_NBR","YEARMONTH")][order(STORE_NBR, YEARMONTH)]
#### Use the functions for calculating correlation 
trial_store <- 88 
corr_nSales <- calculateCorrelation(preTrialMeasures,quote(totSales), trial_store) 
corr_nCustomers <- calculateCorrelation(preTrialMeasures,quote(nCustomers), trial_store)

#### Usethe functions for calculating magnitude
magnitude_nSales <- calculateMagnitudeDistance(preTrialMeasures, quote(totSales),trial_store) 
magnitude_nCustomers <- calculateMagnitudeDistance(preTrialMeasures,
                                                   quote(nCustomers),trial_store) 
#### Create a combined score composed of correlation and magnitude 
corr_weight <- 0.5 
score_nSales <- merge(corr_nSales,magnitude_nSales,by= c("Store1", "Store2"))[,scoreNSales := corr_measure * corr_weight + mag_measure * (1-corr_weight)] 
score_nCustomers <- merge(corr_nCustomers,magnitude_nCustomers,by= c("Store1","Store2"))[,scoreNCust := corr_measure * corr_weight + mag_measure * (1 - corr_weight)] 
####Combine scores across the drivers
score_Control <- merge(score_nSales,score_nCustomers,by= c("Store1", "Store2")) 
score_Control[,finalControlScore:= scoreNSales * 0.5 + scoreNCust * 0.5] 
####Select control stores based on the highest matching store 
####(closest to 1 but not the store itself,i.e.the second ranked highest store) 
####Select control store for trial store 88 
control_store <- score_Control[Store1==trial_store,][order(-finalControlScore)][2,Store2]
control_store
```

# We’ve now found store 237 to be a suitable control store for trial store88
# check if the drivers are indeed similar in the period before the trial
```{r}
####Visual checks on trends based on the drivers
measureOverTimeSales <- measureOverTime 
pastSales <- measureOverTimeSales[,Store_type:= ifelse(STORE_NBR == trial_store,"Trial", 
                                                       ifelse(STORE_NBR == control_store, "Control","Otherstores")) 
                                  ][,totSales:=mean(totSales),by=c("YEARMONTH", "Store_type") 
                                    ][,TransactionMonth:=as.Date(paste(YEARMONTH%/% 100,YEARMONTH %% 100,1,sep="-"),"%Y-%m-%d") 
                                      ][YEARMONTH < 201903,]
ggplot(pastSales, aes(TransactionMonth,totSales,color=Store_type)) + geom_line(aes(linetype=Store_type)) + labs(x= "Month of operation", y= "Total sales", title="Total sales by month")
```
# the trial and control stores have similar total sales

```{r}
####Visual checks on trends based on the drivers 
measureOverTimeCusts <- measureOverTime 
pastCustomers<- measureOverTimeCusts[,Store_type:= ifelse(STORE_NBR == trial_store,"Trial",
                                                          ifelse(STORE_NBR == control_store, "Control","Otherstores")) 
                                     ][,numberCustomers := mean(nCustomers),by= c("YEARMONTH","Store_type") 
                                       ][,TransactionMonth:=as.Date(paste(YEARMONTH%/% 100,YEARMONTH %% 100,1,sep="-"),"%Y-%m-%d")
                                         ][YEARMONTH < 201903,] 

ggplot(pastCustomers, aes(TransactionMonth,numberCustomers,color= Store_type)) + geom_line() + labs(x="Month of operation", y="Total number of customers", title="Total number of customers by month")
```

```{r}
#### Scale pre‐trial control sales to match pre‐tria ltrial store sales
scalingFactorForControlSales <- preTrialMeasures[STORE_NBR== trial_store & YEARMONTH < 201902, sum(totSales)]/preTrialMeasures[STORE_NBR== control_store & YEARMONTH < 201902, sum(totSales)] 
#### Apply the scaling factor 
measureOverTimeSales <- measureOverTime 
scaledControlSales <- measureOverTimeSales[STORE_NBR== control_store,][, controlSales := totSales * scalingFactorForControlSales] 
####Calculate the percentage difference between scaled control sales and trial sales ↪
percentageDiff <- merge(scaledControlSales[,c("YEARMONTH","controlSales")], measureOverTime[STORE_NBR== trial_store, c("totSales", "YEARMONTH")],
                        by= "YEARMONTH" )[,percentageDiff := abs(controlSales-totSales)/controlSales]
####As our null hypothesis is that the trial period is the same as the pre‐tria lperiod, let'stake the standard deviation based on the scaled percentage difference in the pre‐tria lperiod 
stdDev<- sd(percentageDiff[YEARMONTH< 201902,percentageDiff])

degreesOfFreedom<- 7 
####Trial and control store total sales 
measureOverTimeSales <- measureOverTime 
pastSales <- measureOverTimeSales[,Store_type:= ifelse(STORE_NBR == trial_store,"Trial", 
                                                       ifelse(STORE_NBR == control_store, "Control","Otherstores")) 
                                  ][,totSales:=mean(totSales),by=c("YEARMONTH", "Store_type") 
                                    ][,TransactionMonth:=as.Date(paste(YEARMONTH%/% 100,YEARMONTH %% 100,1,sep="-"),"%Y-%m-%d") 
                                      ][Store_type %in% c("Trial","Control"),]
####Control store 95th percentile 
pastSales_Controls95<- pastSales[Store_type== "Control", ][,totSales := totSales * (1 + stdDev * 2) ][,Store_type := "Control 95th% confidence interval"]
####Control store 5th percentile 
pastSales_Controls5<- pastSales[Store_type== "Control", ][,totSales := totSales * (1-stdDev * 2) ][,Store_type := "Control5th%confidence interval"]
trialAssessment <- rbind(pastSales,pastSales_Controls95,pastSales_Controls5) 



```


```{r}
####Plotting these in one nice graph
ggplot(trialAssessment,aes(TransactionMonth,totSales,color=Store_type)) + geom_rect(data=trialAssessment[YEARMONTH< 201905 & YEARMONTH > 201901,], aes(xmin= min(TransactionMonth),xmax= max(TransactionMonth), ymin=0, ymax=Inf,color=NULL),show.legend=FALSE) + 
  geom_line(aes(linetype=Store_type)) + labs(x= "Monthofoperation",y= "Total sales", title="Total sales by month")
```
# The results show that the trial in store88 is significantly different to its control store in the trial period as the trial store performance lies outside of the 5% to 95% confidence interval of the control store in two of the three trial months.
# Now assessing this for number of customers as well

```{r}
####Scale pre‐trial control customers to match pre‐trial trial store customers
scalingFactorForControlCust <- preTrialMeasures[STORE_NBR== trial_store & YEARMONTH < 201902, sum(nCustomers)]/preTrialMeasures[STORE_NBR== control_store & YEARMONTH < 201902, sum(nCustomers)]
####Apply the scaling factor 
measureOverTimeCusts <- measureOverTime 
scaledControlCustomers <- measureOverTimeCusts[STORE_NBR== control_store, ][,controlCustomers:= nCustomers * scalingFactorForControlCust                                                    ][,Store_type:=ifelse(STORE_NBR == trial_store,"Trial", 
                                      ifelse(STORE_NBR == control_store, "Control","Otherstores"))  ]
####Calculate the percentage difference between scaled control sales and trial sales 
percentageDiff <- merge(scaledControlCustomers[,c("YEARMONTH", "controlCustomers")],
measureOverTime[STORE_NBR == trial_store, c("nCustomers", "YEARMONTH")], by = "YEARMONTH" )[, percentageDiff := abs(controlCustomers-nCustomers)/controlCustomers]

#### As our null hypothesis is that the trial period is the same as the pre‐trial period, let's take the standard deviation based on the scaled percentage difference in the pre‐trial period
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff]) 
degreesOfFreedom <- 7
# note that there are 8 months in the pre‐trial period hence 8‐ 1 = 7 degrees of freedom
#### Trial and control store number of customers 
pastCustomers <- measureOverTimeCusts[, nCusts := mean(nCustomers), by = c("YEARMONTH", "Store_type") ][Store_type %in% c("Trial", "Control"), ] 
#### Control store 95th percentile 
pastCustomers_Controls95 <- pastCustomers[Store_type == "Control", ][, nCusts := nCusts * (1 + stdDev * 2) ][, Store_type := "Control 95th % confidence interval"] 
#### Control store 5th percentile 
pastCustomers_Controls5 <- pastCustomers[Store_type == "Control", ][, nCusts := nCusts * (1- stdDev * 2) ][, Store_type := "Control 5th % confidence interval"] 
trialAssessment <- rbind(pastCustomers, pastCustomers_Controls95, pastCustomers_Controls5)

```

```{r}
#### Plotting these in one nice graph 
ggplot(trialAssessment, aes(TransactionMonth, nCusts, color = Store_type)) + geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,], aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 , ymax = Inf, color = NULL), show.legend = FALSE) + 
  geom_line() + labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")
```










