---
title: "Task1 "
output: html_document
date: "2026-01-17"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library(tidyverse)
library(plotly)
library(dlookr)
library(lubridate)
library(dplyr)
library(stringr)
library(xray)
library(data.table)
library(readr)  # for parse_number()
library(scales)
library(ggplot2)
library(ggmosaic)



#### assign the data files to data.tables

transactionData <-  fread(paste0(filePath,"QVI_transaction.csv"))
customerData <-  fread(paste0(filePath,"QVI_customer.csv"))

# EDA. Examine transaction data
```{r}
str(transactionData)
```
```{r}
#### Convert DATE column to a date format
#### A quick Google search tells us that CSV and Excel integer dates begin on 30 Dec 1899
transactionData$DATE <- as.Date(transactionData$DATE, origin = "1899-12-30")
```

# We want to check chips items. We can do some basic text analysis by summarising the individual words in the product name
```{r}
#### Examine the words in PROD_NAME to see if there are any incorrect entries
#### such as products that are not chips
productWords <- data.table(unlist(strsplit(unique(transactionData[,                      PROD_NAME]), " ")))
setnames(productWords, 'words')
```

#As we are only interested in words that will tell us if the product is chips or not, let’s remove all words with digits and special characters such as ‘&’ from our set of product words. We can do this using grepl().
```{r}
#### Removing digits
productWords <- productWords[grepl("\\d", words) == FALSE, ]
#### Removing special characters
productWords <- productWords[grepl("[:alpha:]", words), ]
#### Let's look at the most common words by counting the number of times a word appears and
#### sorting them by this frequency in order of highest to lowest frequency
productWords[, .N, words][order(N, decreasing = TRUE)]

#### Note that sorting by negative N gives us the same result
#productWords[, .N, words][order(-N)]
```

#remove salsa product
```{r}
transactionData[, SALSA := grepl("salsa", tolower(PROD_NAME))]
transactionData <- transactionData[SALSA == FALSE, ][, SALSA := NULL]

```

# use summary() to check summary statistics such as mean, min and max values for each feature to see if there are any obvious outliers in the data and if there are any nulls in any of the columns
```{r}
summary(transactionData)

```
# Let’s investigate further the case where 200 packets of chips are bought in one transaction.
```{r}
#### Filter the dataset to find the outlier
transactionData[PROD_QTY == 200, ]

```

```{r}
#### Let's see if the customer has had other transactions
transactionData[LYLTY_CARD_NBR == 226000, ]

```

# It looks like this customer has only had the two transactions over the year and is not an ordinary retail customer. The customer might be buying chips for commercial purposes instead. We’ll remove this loyalty card number from further analysis
```{r}
#### Filter out the customer based on the loyalty card number
transactionData <- transactionData[LYLTY_CARD_NBR != 226000, ]

#### Re‐examine transaction data
summary(transactionData)
```
# Now, let’s look at the number of transaction lines over time to see if there are any obvious data issues such as missing data.
```{r}
#### Count the number of transactions by date
transactionData[, .N, by = DATE]
```

#Let’s create a sequence of dates from 1 Jul 2018 to 30 Jun 2019 and use this to create a chart of number of transactions over time to find the missing date
```{r}
#### Create a sequence of dates and join this the count of transactions by date
allDates <- data.table(seq(as.Date("2018/07/01"), as.Date("2019/06/30"), by = "day"))
setnames(allDates, "DATE")
transactions_by_day <- merge(allDates, transactionData[, .N, by = DATE], all.x = TRUE)
#### Setting plot themes to format graphs
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
#### Plot transactions over time
ggplot(transactions_by_day, aes(x = DATE, y = N)) + geom_line() + labs(x = "Day", y = "Number of transactions", title = "Transactions over time") + scale_x_date(breaks = "1 month") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

# there is an increase in purchases in December and a break in late December. Let’s zoom in on this
```{r}
#### Filter to December and look at individual days
ggplot(transactions_by_day[month(DATE) == 12, ], aes(x = DATE, y = N)) + geom_line() +
labs(x = "Day", y = "Number of transactions", title = "Transactions over time") +
scale_x_date(breaks = "1 day") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```
# the increase in sales occurs in the lead-up to Christmas and that there are zero sales on Christmas day itself. This is due to shops being closed on Christmas day.

# Now that we are satisfied that the data no longer has outliers, we can move on to creating other features such as brand of chips or pack size from PROD_NAME. We will start with pack size.
```{r}
#### Pack size
#### We can work this out by taking the digits that are in PROD_NAME
transactionData[, PACK_SIZE := parse_number(PROD_NAME)]
#### Let's check if the pack sizes look sensible
transactionData[, .N, PACK_SIZE][order(PACK_SIZE)]
```

```{r}
#### Let's plot a histogram of PACK_SIZE since we know that it is a categorical variable and not a continuous variable even though it is numeric.
hist(transactionData[, PACK_SIZE])
```

# now to create brands, we can use the first word in PROD_NAME to work out the brand name
```{r}
#### Brands
transactionData[, BRAND := toupper(substr(PROD_NAME, 1, regexpr(pattern = ' ',PROD_NAME) - 1))]
#### Checking brands
transactionData[, .N, by = BRAND][order(-N)]
```

```{r}
#### Clean brand names
transactionData[BRAND == "RED", BRAND := "RRD"]
transactionData[BRAND == "SNBTS", BRAND := "SUNBITES"]
transactionData[BRAND == "INFZNS", BRAND := "INFUZIONS"]
transactionData[BRAND == "WW", BRAND := "WOOLWORTHS"]
transactionData[BRAND == "SMITH", BRAND := "SMITHS"]
transactionData[BRAND == "NCC", BRAND := "NATURAL"]
transactionData[BRAND == "DORITO", BRAND := "DORITOS"]
transactionData[BRAND == "GRAIN", BRAND := "GRNWVES"]
#### Check again
transactionData[, .N, by = BRAND][order(BRAND)]

```
-----------------------------------------------------------

#Examining customer data
```{r}
str(customerData)
```
```{r}
summary(customerData)
```

```{r}
#### Examining the values of lifestage and premium_customer
customerData[, .N, by = LIFESTAGE][order(-N)]

```

```{r}
customerData[, .N, by = PREMIUM_CUSTOMER][order(-N)]
```

#As there do not seem to be any issues with the customer data, we can now go ahead and join the transaction and customer data sets together
```{r}
#### Merge transaction data to customer data
data <- merge(transactionData, customerData, all.x = TRUE)
```

# As the number of rows in data is the same as that of transactionData, we can be sure that no duplicates were created. This is because we created data by setting all.x = TRUE (in other words, a left join) which means take all the rows in transactionData and find rows with matching values in shared columns and then joining the details in these rows to the x or the first mentioned table

```{r}
#Let’s also check if some customers were not matched on by checking for nulls.
data[is.null(LIFESTAGE), .N]
data[is.null(PREMIUM_CUSTOMER), .N]
```
# Data analysis on customer segments
# we can define some metrics of interest to the client:

# which group buy the most of chips, what is their avg money spent
# avg money spent of PREMIUM_CUSTOMER
# buying pattern (time/seasonal?) 


# Who spends the most on chips (total sales), describing customers by lifestage and how premium their general purchasing behaviour is
# How many customers are in each segment
# How many chips are bought per customer by segment
# What’s the average chip price by customer segment

# Let’s start with calculating total sales by LIFESTAGE and PREMIUM_CUSTOMER and plotting the split by these segments to describe which customer segment contribute most to chip sales.
```{r}
#### Total sales by LIFESTAGE and PREMIUM_CUSTOMER
sales <- data[, .(SALES = sum(TOT_SALES)), .(LIFESTAGE, PREMIUM_CUSTOMER)]
#### Create plot
p <- ggplot(data = sales) + geom_mosaic(aes(weight = SALES, x = product(PREMIUM_CUSTOMER, LIFESTAGE),
fill = PREMIUM_CUSTOMER)) + labs(x = "Lifestage", y = "Premium customer flag", title = "Proportion of
sales") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

#### Plot and label with proportion of sales
p + geom_text(data = ggplot_build(p)$data[[1]], aes(x = (xmin + xmax)/2 , y =  (ymin + ymax)/2, label = as.character(paste(round(.wt/sum(.wt),3)*100,'%'))))
```

# Let’s see if the higher sales are due to there being more customers who buy chips
```{r}
#### Number of customers by LIFESTAGE and PREMIUM_CUSTOMER
customers <- data[, .(CUSTOMERS = uniqueN(LYLTY_CARD_NBR)), .(LIFESTAGE, PREMIUM_CUSTOMER)][order(-CUSTOMERS)]
#### Create plot
p <-  ggplot(data = customers) + geom_mosaic(aes(weight = CUSTOMERS, x = product(PREMIUM_CUSTOMER, LIFESTAGE), fill = PREMIUM_CUSTOMER)) + labs(x = "Lifestage", y = "Premium customer flag", title = "Proportion of customers") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

#### Plot and label with proportion of customers
p + geom_text(data = ggplot_build(p)$data[[1]], aes(x = (xmin + xmax)/2 , y = (ymin + ymax)/2, label = as.character(paste(round(.wt/sum(.wt),3)*100, '%'))))
```
# There are more Mainstream - young singles/couples and Mainstream - retirees who buy chips. This contributes to there being more sales to these customer segments but this is not a major driver for the Budget - Older families segment

# Higher sales may also be driven by more units of chips being bought per customer. Let’s have a look at this next.
```{r}
#### Average number of units per customer by LIFESTAGE and PREMIUM_CUSTOMER
avg_units <- data[, .(AVG = sum(PROD_QTY)/uniqueN(LYLTY_CARD_NBR)), .(LIFESTAGE, PREMIUM_CUSTOMER)][order(-AVG)]
#### Create plot
ggplot(data = avg_units, aes(weight = AVG, x = LIFESTAGE, fill = PREMIUM_CUSTOMER)) + geom_bar(position = position_dodge()) + labs(x = "Lifestage", y = "Avg units per transaction", title = "Units per
customer") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```
#Older families and young families in general buy more chips per customer
#Let’s also investigate the average price per unit chips bought for each customer segment as this is also a driver of total sales.

```{r}
#### Average price per unit by LIFESTAGE and PREMIUM_CUSTOMER
avg_price <- data[, .(AVG = sum(TOT_SALES)/sum(PROD_QTY)), .(LIFESTAGE, PREMIUM_CUSTOMER)][order(-AVG)]
#### Create plot
ggplot(data = avg_price, aes(weight = AVG, x = LIFESTAGE, fill = PREMIUM_CUSTOMER)) + geom_bar(position = position_dodge()) + labs(x = "Lifestage", y = "Avg price per unit", title = "Price per unit") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

# Mainstream midage and young singles and couples are more willing to pay more per packet of chips compared to their budget and premium counterparts. This may be due to premium shoppers being more likely to buy healthy snacks and when they buy chips, this is mainly for entertainment purposes rather than their own consumption. This is also supported by there being fewer premium midage and young singles and couples buying chips compared to their mainstream counterparts.

#As the difference in average price per unit isn’t large, we can check if this difference is statistically different. 

```{r}
#### Perform an independent t‐test between mainstream vs premium and budget midage and
#### young singles and couples
pricePerUnit <- data[, price := TOT_SALES/PROD_QTY]
t.test(data[LIFESTAGE %in% c("YOUNG SINGLES/COUPLES", "MIDAGE SINGLES/COUPLES") & PREMIUM_CUSTOMER == "Mainstream", price] ,
       data[LIFESTAGE %in% c("YOUNG SINGLES/COUPLES", "MIDAGE SINGLES/COUPLES") & PREMIUM_CUSTOMER != "Mainstream", price] , alternative = "greater")
```
# The t-test results in a p-value < 2.2e-16, i.e. the unit price for mainstream, young and mid-age singles and couples are significantly higher than that of budget or premium, young and midage singles and couples.

-------------------------------------------------------------
#Deep dive into specific customer segments for insights
#We might want to target customer segments that contribute the most to sales to retain them or further increase sales. Let’s look at Mainstream - young singles/couples. For instance, let’s find out if they tend to buy a particular brand of chips.

```{r}
#### Deep dive into Mainstream, young singles/couples
segment1 <- data[LIFESTAGE == "YOUNG SINGLES/COUPLES" & PREMIUM_CUSTOMER == "Mainstream",]
other <- data[!(LIFESTAGE == "YOUNG SINGLES/COUPLES" & PREMIUM_CUSTOMER == "Mainstream"),]
#### Brand affinity compared to the rest of the population
quantity_segment1 <- segment1[, sum(PROD_QTY)]
quantity_other <- other[, sum(PROD_QTY)]
quantity_segment1_by_brand <- segment1[, .(targetSegment = sum(PROD_QTY)/quantity_segment1), by = BRAND]
quantity_other_by_brand <- other[, .(other = sum(PROD_QTY)/quantity_other), by = BRAND]
brand_proportions <- merge(quantity_segment1_by_brand, quantity_other_by_brand)[, affinityToBrand := targetSegment/other]
brand_proportions[order(-affinityToBrand)]

```

# We can see that :
#Mainstream young singles/couples are 23% more likely to purchase Tyrrells chips compared to the rest of the population
# Mainstream young singles/couples are 56% less likely to purchase Burger Rings compared to the rest of the population

#Let’s also find out if our target segment tends to buy larger packs of chips.
```{r}
#### Preferred pack size compared to the rest of the population
quantity_segment1_by_pack <- segment1[, .(targetSegment = sum(PROD_QTY)/quantity_segment1), by = PACK_SIZE]
quantity_other_by_pack <- other[, .(other = sum(PROD_QTY)/quantity_other), by = PACK_SIZE]
pack_proportions <- merge(quantity_segment1_by_pack, quantity_other_by_pack)[, affinityToPack := targetSegment/other]
pack_proportions[order(-affinityToPack)]

```

# It looks like Mainstream young singles/couples are 27% more likely to purchase a 270g pack of chips compared to the rest of the population but let’s dive into what brands sell this pack size.
```{r}
data[PACK_SIZE == 270, unique(PROD_NAME)]
```
#Twisties are the only brand offering 270g packs and so this may instead be reflecting a higher likelihood of purchasing Twisties.










